---
title: "LAB1"
author: "Evelyn Zhang"
date: "9/19/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(ggplot2)
```


## Read data

```{r}
data<-read.csv('data/boston_data.csv')
data
```

## Plots and regression with each variable against y

```{r}
x<- data$CRIM
y<- data$MEDV

plot(x, y
     ,xlab="Per capita crime rate by town"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)

minX <- min(x)
maxX <- max(x)

#linear regression
simpleLinearRegression <- lm(y~x,data=data)
summary(simpleLinearRegression)
#quadratic regression
quadraticRegression <- lm(y~x+I(x^2),data=data)
summary(quadraticRegression)
#cubic regression
cubicRegression <- lm(y~x+I(x^2)+I(x^3),data=data)
summary(cubicRegression)
#inverse x regression
inverseRegression <- lm(y~I(1/x),data=data)
summary(inverseRegression)

#linear regression
betas <- coef(simpleLinearRegression)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(minX,maxX,0.01)
linearYPredictions <- beta0 + beta1*xVals
lines(xVals,linearYPredictions,col='red')

ssr_SLR = sum((fitted(simpleLinearRegression) - y)^2) #NEW

#quadratic regression
betas <- coef(quadraticRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX,0.01)
quadraticYPredictions <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals,quadraticYPredictions,col='blue', lty=2)

ssr_QR = sum((fitted(quadraticRegression) - y)^2) #NEW

#cubic regression
betas <- coef(cubicRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]
beta3 <- betas[4]

xVals <- seq(minX,maxX,0.01)
cubicYPredictions <- beta0 + beta1*xVals + beta2*xVals^2 + beta3*xVals^3
lines(xVals,cubicYPredictions,col='orange',lty=5)

ssr_CR = sum((fitted(cubicRegression) - y)^2) #NEW

sprintf("SSR Simple Linear Regression = %.2f",ssr_SLR)
sprintf("SSR Quadratic Regression = %.2f",ssr_QR)
sprintf("SSR Cubic Regression = %.2f",ssr_CR)

#inverse regression
betas <- coef(inverseRegression)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(minX,maxX,0.01)
#***How do I fit a y=1/x regression? 
linearYPredictions <- beta0 + beta1/xVals
lines(xVals,linearYPredictions,col='green')


legend(1,26,legend=c("Lin Reg.","Quad. Reg", "Cubic Reg.", "Inverse Reg." )
                     ,col=c('red','blue','orange', 'green')
                     ,lty=c(1,2,5)
)

```
```{r}
plot(data$ZN, data$MEDV
     ,xlab="Proportion of residential land zoned for lots over 25,000 sq. ft"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)
```
```{r}
x<- data$INDUS
y<- data$MEDV

plot(x, y
     ,xlab="Proportion of non-retail business acres per town"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)

minX <- min(x)
maxX <- max(x)

#linear regression
simpleLinearRegression <- lm(y~x,data=data)
summary(simpleLinearRegression)
#quadratic regression
quadraticRegression <- lm(y~x+I(x^2),data=data)
summary(quadraticRegression)
#cubic regression
cubicRegression <- lm(y~x+I(x^2)+I(x^3),data=data)
summary(cubicRegression)

#linear regression
betas <- coef(simpleLinearRegression)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(minX,maxX,0.01)
linearYPredictions <- beta0 + beta1*xVals
lines(xVals,linearYPredictions,col='red')

ssr_SLR = sum((fitted(simpleLinearRegression) - y)^2) #NEW

#quadratic regression
betas <- coef(quadraticRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX,0.01)
quadraticYPredictions <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals,quadraticYPredictions,col='blue', lty=2)

ssr_QR = sum((fitted(quadraticRegression) - y)^2) #NEW


#cubic regression
betas <- coef(cubicRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]
beta3 <- betas[4]

xVals <- seq(minX,maxX,0.01)
cubicYPredictions <- beta0 + beta1*xVals + beta2*xVals^2 + beta3*xVals^3
lines(xVals,cubicYPredictions,col='orange',lty=5)

ssr_CR = sum((fitted(cubicRegression) - y)^2) #NEW

sprintf("SSR Simple Linear Regression = %.2f",ssr_SLR)
sprintf("SSR Quadratic Regression = %.2f",ssr_QR)
sprintf("SSR Cubic Regression = %.2f",ssr_CR)


legend(1,26,legend=c("Lin Reg.","Quad. Reg", "Cubic Reg." )
                     ,col=c('red','blue','orange')
                     ,lty=c(1,2,5)
)


```

```{r}
x<- data$CHAS
y<- data$MEDV

plot(x, y
     ,xlab="Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)
```


```{r}
x<- data$NOX
y<- data$MEDV

plot(x, y
     ,xlab="Nitric oxide concentration (parts per 10 million)"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)

minX <- min(x)
maxX <- max(x)

#linear regression
simpleLinearRegression <- lm(y~x,data=data)
summary(simpleLinearRegression)
#quadratic regression
quadraticRegression <- lm(y~x+I(x^2),data=data)
summary(quadraticRegression)
#cubic regression
cubicRegression <- lm(y~x+I(x^2)+I(x^3),data=data)
summary(cubicRegression)

#linear regression
betas <- coef(simpleLinearRegression)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(minX,maxX,0.01)
linearYPredictions <- beta0 + beta1*xVals
lines(xVals,linearYPredictions,col='red')

ssr_SLR = sum((fitted(simpleLinearRegression) - y)^2) #NEW

#quadratic regression
betas <- coef(quadraticRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX,0.01)
quadraticYPredictions <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals,quadraticYPredictions,col='blue', lty=2)

ssr_QR = sum((fitted(quadraticRegression) - y)^2) #NEW

#cubic regression
betas <- coef(cubicRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]
beta3 <- betas[4]

xVals <- seq(minX,maxX,0.01)
cubicYPredictions <- beta0 + beta1*xVals + beta2*xVals^2 + beta3*xVals^3
lines(xVals,cubicYPredictions,col='orange',lty=5)

ssr_CR = sum((fitted(cubicRegression) - y)^2) #NEW

sprintf("SSR Simple Linear Regression = %.2f",ssr_SLR)
sprintf("SSR Quadratic Regression = %.2f",ssr_QR)
sprintf("SSR Cubic Regression = %.2f",ssr_CR)

legend(1,26,legend=c("Lin Reg.","Quad. Reg", "Cubic Reg." )
                     ,col=c('red','blue','orange')
                     ,lty=c(1,2,5)
)

```

```{r}
x<- data$RM
y<- data$MEDV

plot(x, y
     ,xlab="Average number of rooms per dwelling"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)

#linear regression
simpleLinearRegression <- lm(y~x,data=data)
summary(simpleLinearRegression)
#quadratic regression
quadraticRegression <- lm(y~x+I(x^2),data=data)
summary(quadraticRegression)
#cubic regression
cubicRegression <- lm(y~x+I(x^2)+I(x^3),data=data)
summary(cubicRegression)

minX <- min(x)
maxX <- max(x)

#linear regression
betas <- coef(simpleLinearRegression)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(minX,maxX,0.01)
linearYPredictions <- beta0 + beta1*xVals
lines(xVals,linearYPredictions,col='red')

ssr_SLR = sum((fitted(simpleLinearRegression) - y)^2) #NEW

#quadratic regression
betas <- coef(quadraticRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX,0.01)
quadraticYPredictions <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals,quadraticYPredictions,col='blue', lty=2)

ssr_QR = sum((fitted(quadraticRegression) - y)^2) #NEW

#cubic regression
betas <- coef(cubicRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]
beta3 <- betas[4]

xVals <- seq(minX,maxX,0.01)
cubicYPredictions <- beta0 + beta1*xVals + beta2*xVals^2 + beta3*xVals^3
lines(xVals,cubicYPredictions,col='orange',lty=5)

ssr_CR = sum((fitted(cubicRegression) - y)^2) #NEW

sprintf("SSR Simple Linear Regression = %.2f",ssr_SLR)
sprintf("SSR Quadratic Regression = %.2f",ssr_QR)
sprintf("SSR Cubic Regression = %.2f",ssr_CR)


legend(1,26,legend=c("Lin Reg.","Quad. Reg", "Cubic Reg." )
                     ,col=c('red','blue','orange')
                     ,lty=c(1,2,5)
)

```

```{r}
x<- data$AGE
y<- data$MEDV

plot(x, y
     ,xlab="Proportion of owner-occupied units built prior to 1940"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)

#linear regression
simpleLinearRegression <- lm(y~x,data=data)
summary(simpleLinearRegression)
#quadratic regression
quadraticRegression <- lm(y~x+I(x^2),data=data)
summary(quadraticRegression)
#cubic regression
cubicRegression <- lm(y~x+I(x^2)+I(x^3),data=data)
summary(cubicRegression)

minX <- min(x)
maxX <- max(x)

#linear regression
betas <- coef(simpleLinearRegression)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(minX,maxX,0.01)
linearYPredictions <- beta0 + beta1*xVals
lines(xVals,linearYPredictions,col='red')

ssr_SLR = sum((fitted(simpleLinearRegression) - y)^2) #NEW

#quadratic regression
betas <- coef(quadraticRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX,0.01)
quadraticYPredictions <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals,quadraticYPredictions,col='blue', lty=2)

ssr_QR = sum((fitted(quadraticRegression) - y)^2) #NEW

#cubic regression
betas <- coef(cubicRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]
beta3 <- betas[4]

xVals <- seq(minX,maxX,0.01)
cubicYPredictions <- beta0 + beta1*xVals + beta2*xVals^2 + beta3*xVals^3
lines(xVals,cubicYPredictions,col='orange',lty=5)

ssr_CR = sum((fitted(cubicRegression) - y)^2) #NEW

sprintf("SSR Simple Linear Regression = %.2f",ssr_SLR)
sprintf("SSR Quadratic Regression = %.2f",ssr_QR)
sprintf("SSR Cubic Regression = %.2f",ssr_CR)


legend(1,26,legend=c("Lin Reg.","Quad. Reg", "Cubic Reg." )
                     ,col=c('red','blue','orange')
                     ,lty=c(1,2,5)
)


```

```{r}
x<- data$DIS
y<- data$MEDV

plot(x, y
     ,xlab="Weighted distances to five Boston employment centers"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)

#linear regression
simpleLinearRegression <- lm(y~x,data=data)
summary(simpleLinearRegression)
#quadratic regression
quadraticRegression <- lm(y~x+I(x^2),data=data)
summary(quadraticRegression)
#cubic regression
cubicRegression <- lm(y~x+I(x^2)+I(x^3),data=data)
summary(cubicRegression)

minX <- min(x)
maxX <- max(x)

#linear regression
betas <- coef(simpleLinearRegression)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(minX,maxX,0.01)
linearYPredictions <- beta0 + beta1*xVals
lines(xVals,linearYPredictions,col='red')

ssr_SLR = sum((fitted(simpleLinearRegression) - y)^2) #NEW

#quadratic regression
betas <- coef(quadraticRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX,0.01)
quadraticYPredictions <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals,quadraticYPredictions,col='blue', lty=2)

ssr_QR = sum((fitted(quadraticRegression) - y)^2) #NEW

#cubic regression
betas <- coef(cubicRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]
beta3 <- betas[4]

xVals <- seq(minX,maxX,0.01)
cubicYPredictions <- beta0 + beta1*xVals + beta2*xVals^2 + beta3*xVals^3
lines(xVals,cubicYPredictions,col='orange',lty=5)

ssr_CR = sum((fitted(cubicRegression) - y)^2) #NEW

sprintf("SSR Simple Linear Regression = %.2f",ssr_SLR)
sprintf("SSR Quadratic Regression = %.2f",ssr_QR)
sprintf("SSR Cubic Regression = %.2f",ssr_CR)


legend(1,26,legend=c("Lin Reg.","Quad. Reg", "Cubic Reg." )
                     ,col=c('red','blue','orange')
                     ,lty=c(1,2,5)
       
       
)

```



```{r}
x<- data$RAD
y<- data$MEDV

plot(x, y
     ,xlab="Index of accessibility to radial highways"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)

```


```{r}
x<- data$TAX
y<- data$MEDV

plot(x, y
     ,xlab="Full-value property tax rate per $10,000"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)

```


```{r}
x<- data$B
y<- data$MEDV

plot(x, y
     ,xlab="1000(Bk - 0.63)², where Bk is the proportion of [people of African American descent] by town"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)

```


```{r}
x<- data$LSTAT
y<- data$MEDV

plot(x, y
     ,xlab="Percentage of lower status of the population"
     ,ylab="Median value of owner-occupied homes in $1000s"
     ,tck=0.02)

#linear regression
simpleLinearRegression <- lm(y~x,data=data)
summary(simpleLinearRegression)
#quadratic regression
quadraticRegression <- lm(y~x+I(x^2),data=data)
summary(quadraticRegression)
#cubic regression
cubicRegression <- lm(y~x+I(x^2)+I(x^3),data=data)
summary(cubicRegression)

minX <- min(x)
maxX <- max(x)

#linear regression
betas <- coef(simpleLinearRegression)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(minX,maxX,0.01)
linearYPredictions <- beta0 + beta1*xVals
lines(xVals,linearYPredictions,col='red')

ssr_SLR = sum((fitted(simpleLinearRegression) - y)^2) #NEW

#quadratic regression
betas <- coef(quadraticRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX,0.01)
quadraticYPredictions <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals,quadraticYPredictions,col='blue', lty=2)

ssr_QR = sum((fitted(quadraticRegression) - y)^2) #NEW

#cubic regression
betas <- coef(cubicRegression)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]
beta3 <- betas[4]

xVals <- seq(minX,maxX,0.01)
cubicYPredictions <- beta0 + beta1*xVals + beta2*xVals^2 + beta3*xVals^3
lines(xVals,cubicYPredictions,col='orange',lty=5)

ssr_CR = sum((fitted(cubicRegression) - y)^2) #NEW

sprintf("SSR Simple Linear Regression = %.2f",ssr_SLR)
sprintf("SSR Quadratic Regression = %.2f",ssr_QR)
sprintf("SSR Cubic Regression = %.2f",ssr_CR)

legend(1,26,legend=c("Lin Reg.","Quad. Reg", "Cubic Reg." )
                     ,col=c('red','blue','orange')
                     ,lty=c(1,2,5)
)


```


# Put all relevant variables together, it seems that cubic model is the most accurate according to SSR
```{r}
x1<- data$CRIM
x2<- data$INDUS
x3<- data$NOX
x4<- data$RM
x5<- data$AGE
x6<- data$DIS
x7<- data$LSTAT
y<- data$MEDV

totalRegression <- lm(y~x1+x2+x3+x4+x5+x6+x7+I(x1^2)+I(x2^2)+I(x3^2)+I(x4^2)+I(x1^5)+I(x6^2)+I(x7^2)+I(x1^3)+I(x2^3)+I(x3^3)+I(x4^3)+I(x5^3)+I(x6^3)+I(x7^3),data=data) 

summary(totalRegression)

```

### Questions

# 1. How would you explain SSE?
# SSE stands for sum of squared errors. It is the sum of the squared difference between predicted y values using x variables and the actual y values from the sample. 

# 2. How would you explain your model?
# By looking at the plot of each x-variable against y, I determined that CRIM, INDUS, NOX, RM, AGE, DIS, and LSTAT are the x-variables related to y. By looking at SSR of SLR, QR and CR, the best model seems to be the cubic model. Thus, the final regression is a cubic regressio of MEDV on the seven factors. 

### Additional Questions for the assignment
# 1. How can I best determine which variables are related to y? 
# 2. How to fit a y=1/x regression or y=ln(x) regression? 
# 3. How do I put polynomial and multiple regerssion together? 
# 4. How to deal with discrete variables? Ex. RAD: Index of accessibility to radial highways, CHAS: Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)

