---
title: "Stephanie Spitzer Simple Multiple Polynomial Lab"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(tidyverse)
library(ggplot2)

boston_data <- read.csv('data/boston_data.csv')
boston_data

slr <- lm(formula = PTRATIO ~ TAX, data = boston_data)
summary(slr)

mlr <- lm(formula = PTRATIO ~ CRIM + RM + DIS + TAX + LSTAT, data = boston_data)
summary(mlr)

plr <- lm(formula = PTRATIO ~ poly(TAX, 3), data = boston_data)
summary(plr)

ggplot(data = boston_data, mapping = aes(x = TAX, y = PTRATIO)) + 
  geom_point() +
  geom_smooth(method = "lm")
```

##Questions

1. How would you explain SSE?

The SSE is the sum of squared difference between each observation and the mean. A SSE value closer to 0 indicates that the model has a small error component and the model fits the data more closely.

2. How would you explain your model?
For the simple linear regression, for every $1 increase in the property tax per $10,000, the price increases by $0.00592. For the multiple linear regression, a 1% increase in the crime rate will increase the price by $0.0029466, a one room increase in the number of rooms in a house will decrease the price by $0.6830991, a 1 mile increase in the distance from five Boston employment centers increases the price by $0.0497357, a $1 increase in the property tax per $10,000 will increase the price by $0.0050266, and a 1% increase in a population with a lower status will increase the price by $0.0132803. For the polynomial regression, since the data does not follow a linear pattern, we used a third degree polynomial of the TAX variable to help explain the price for houses in Boston. 
