---
title: "Class4Lab"
output: html_document
name: "Yongyi (Sophie) Peng"
---

**The Goal**
Given a data set containing characteristics about the Boston housing market, explore methods for predicting housing prices.

You'll need to clone this repository using RStudio first. 
Practice writing and adding code that produces a simple linear regression, multiple linear regression, and polynomial regression.
Include a visual and quantitative method for assessing whether some models may better explain housing prices than other models.

We will discuss some models students have proposed and ways to communicate these models to a lay audience.

### Data
The dataset and the data dictionary are included in the folder called data.

### Questions

* How would you explain SSE?
* How would you explain your model?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
Housing <- read.csv('data/boston_data_updated.csv')
head(Housing)
summary(Housing)
```




##(1) Multiple linear regression model:
Y is MEDV which represents median value of owner-occupied homes in $1000s
```{r}
multipleLinearRegressionOne <- lm(MEDV ~ CRIM+ZN+INDUS+CHAS+NOX+RM+AGE+DIS+RAD+TAX+PTRATIO+B+LSTAT, data = Housing)

print(multipleLinearRegressionOne)

```

The predicted model is :

predicted MEDV = 3.646e+01 -1.080e-01*CRIM + 4.642e-02*ZN+2.056e-02*INDUS+2.687e+00*CHAS-1.777e+01*NOX+3.810e+00*RM+6.922e-04*AGE-1.476e+00*DIS+3.060e-01*RAD-1.233e-02*TAX-9.527e-01*PTRATIO+9.312e-03*B-5.248e-01*LSTAT

```{r pressure, echo=FALSE}
plot(Housing)
```



```{r pressure, echo=FALSE}
plot(multipleLinearRegressionOne)
```

From the QQ plot, we can see that the right tail doesn't fit the line very well, but overall not too bad.

```{r}
summary(multipleLinearRegressionOne)
confint(multipleLinearRegressionOne)
```
The R-squared value is pretty good which means that the model fit the data pretty well.


###Only use four predictors to compare the models:
In this part, I only used four predictors instead of using all the predictors. I will use predictors CRIM, RM, INDUS, and DIS in simple linear regression and polynomial regression model. 

Multiple linear regression:
```{r}
multipleLinearRegressionTwo <- lm(MEDV ~ CRIM+RM+INDUS+DIS, data = Housing)
print(multipleLinearRegressionTwo)

```

Model:
predicted MEDV = -17.2026-0.2181*CRIM+7.4761*RM-0.3709*INDUS-0.6142*DIS

```{r pressure, echo=FALSE}
plot(multipleLinearRegressionTwo)
```
```{r}
summary(multipleLinearRegressionTwo)
confint(multipleLinearRegressionTwo)
```
The adjusted R-squared value decrease after the number of predictors decrease.
Since I only used CRIM, RM, INDUS, and DIS as predictors in simple linear regression and polynomial regression, I would only look at the second multiple linear regression model which is predicted MEDV = -17.2026-0.2181*CRIM+7.4761*RM-0.3709*INDUS-0.6142*DIS.






##(2) Simple linear Regression:
For simple linear regression, I will only try four predictors since there are so many predictors in the dataset.

```{r}
head(Housing)
```

####The first predictor: CRIM
The first predictor I chose is CRIM which is the the per capita crime rate by town.
Model: predicted MEDV = 24.0331-0.4152*CRIM
```{r}
simpleLinearRegressionOne <- lm(MEDV~CRIM, data = Housing)
print(simpleLinearRegressionOne)
```
```{r pressure, echo=FALSE}

plot(Housing$CRIM, Housing$MEDV, xlab = "CRIM", ylab="MEDV")
betas <- coef(simpleLinearRegressionOne)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(min(Housing$CRIM), max(Housing$CRIM),1)
yPredictions <- beta0 + beta1*xVals
lines(xVals, yPredictions, col='red')

```
The scatterplot shows that the linear relationship between MEDV and CRIM is very weak.Thus, I don't think simple linear regression is a good model for this data. 

```{r}
summary(simpleLinearRegressionOne)
confint(simpleLinearRegressionOne)
```
The R-squared value is weak. This indicates that the model doesn't fit the data very well.



####The second predictor: RM
The second predictor I chose is RM which is the average number of rooms per dwelling
Model: predicted MEDV =-34.671+9.102*RM         
```{r}
simpleLinearRegressionTwo <- lm(MEDV~RM, data = Housing)
print(simpleLinearRegressionTwo)
```
```{r pressure, echo=FALSE}

plot(Housing$RM, Housing$MEDV, xlab = "RM", ylab="MEDV")
betas <- coef(simpleLinearRegressionTwo)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(min(Housing$RM), max(Housing$RM),1)
yPredictions <- beta0 + beta1*xVals
lines(xVals, yPredictions, col='red')

```
From the scatterplot, we can see that there's a linear relationship between RM and MEDV.

```{r}
summary(simpleLinearRegressionTwo)
confint(simpleLinearRegressionTwo)
```
The R-squared value is weak, this indicates that the model doesn't fit the data very well.



####The third predictor: INDUS
The third predictor I chose is INDUS which is the proportion of non-retail business acres per town.
Model: predicted MEDV = 29.7549-0.6485*INDUS
```{r}
simpleLinearRegressionThree <- lm(MEDV~INDUS, data = Housing)
print(simpleLinearRegressionThree)
```
```{r pressure, echo=FALSE}

plot(Housing$INDUS, Housing$MEDV, xlab = "INDUS", ylab="MEDV")

betas <- coef(simpleLinearRegressionThree)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(min(Housing$INDUS), max(Housing$INDUS),1)
yPredictions <- beta0 + beta1*xVals
lines(xVals, yPredictions, col='red')

```
From the scatterplot, I can see a very weak negative linear relationship between INDUS and MEDV.

```{r}
summary(simpleLinearRegressionThree)
confint(simpleLinearRegressionThree)
```
The R-squared value is very weak, this indicates that the model doesn't fit the data very well.






####The fourth predictor: DIS
The fourth predictor I chose is DIS which is the weighted distances to five Boston employment centers
Model: predicted MEDV = 18.390+1.092*DIS
```{r}
simpleLinearRegressionFour <- lm(MEDV~DIS, data = Housing)
print(simpleLinearRegressionFour)
```
```{r pressure, echo=FALSE}

plot(Housing$DIS, Housing$MEDV, xlab = "DIS", ylab="MEDV")

betas <- coef(simpleLinearRegressionFour)
beta0 <- betas[1]
beta1 <- betas[2]

xVals <- seq(min(Housing$DIS), max(Housing$DIS),1)
yPredictions <- beta0 + beta1*xVals
lines(xVals, yPredictions, col='red')

```
It is hard to tell whether there's a linear relationship between DIS and MEDV.

In conclude, after using four predictors to do simple linear regression, I don't think simple linear regression is the best model to do prediction.

```{r}
summary(simpleLinearRegressionFour)
confint(simpleLinearRegressionFour)
```
The R-squared value is super weak, this indicates that the model doesn't fit the data.







##(3) Polynomial Regression:

####The first predictor is CRIM
```{r}
quadraticRegressionOne <- lm(MEDV~CRIM+I(CRIM^2), data=Housing)
print(quadraticRegressionOne)
```

```{r pressure, echo=FALSE}
plot(Housing$CRIM, Housing$MEDV, xlab="CRIM", ylab = "MEDV")

minX <- min(Housing$CRIM)
maxX <- max(Housing$CRIM)

betas <- coef(quadraticRegressionOne)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX, 0.01)
quadraticYPredictionOne <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals, quadraticYPredictionOne, col = 'blue', lty=2)

```
```{r}
summary(quadraticRegressionOne)
confint(quadraticRegressionOne)
```
The R-squared value is small, which indicates that the model doesn't fit the data very well.



####The second predictor is RM
```{r}
quadraticRegressionTwo <- lm(MEDV~RM+I(RM^2), data=Housing)
print(quadraticRegressionTwo)
```
```{r pressure, echo=FALSE}
plot(Housing$RM, Housing$MEDV, xlab="RM", ylab = "MEDV")

minX <- min(Housing$RM)
maxX <- max(Housing$RM)

betas <- coef(quadraticRegressionTwo)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX, 0.01)
quadraticYPredictionTwo <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals, quadraticYPredictionTwo, col = 'blue', lty=2)

```
```{r}
summary(quadraticRegressionTwo)
confint(quadraticRegressionTwo)
```
The R-squared value here is fine. Not too bad.





####The third predictor is INDUS
```{r}
quadraticRegressionThree <- lm(MEDV~INDUS+I(INDUS^2), data=Housing)
print(quadraticRegressionThree)
```
```{r pressure, echo=FALSE}
plot(Housing$INDUS, Housing$MEDV, xlab="INDUS", ylab = "MEDV")

minX <- min(Housing$INDUS)
maxX <- max(Housing$INDUS)

betas <- coef(quadraticRegressionThree)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX, 0.01)
quadraticYPredictionThree <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals, quadraticYPredictionThree, col = 'blue', lty=2)

```
```{r}
summary(quadraticRegressionThree)
confint(quadraticRegressionThree)
```
The R-squared value here is small.Not good.




####The fourth predictor is DIS
```{r}
quadraticRegressionFourth <- lm(MEDV~DIS+I(DIS^2), data=Housing)
print(quadraticRegressionFourth)
```

```{r pressure, echo=FALSE}
plot(Housing$DIS, Housing$MEDV, xlab="DIS", ylab = "MEDV")

minX <- min(Housing$DIS)
maxX <- max(Housing$DIS)

betas <- coef(quadraticRegressionFourth)
beta0 <- betas[1]
beta1 <- betas[2]
beta2 <- betas[3]

xVals <- seq(minX,maxX, 0.01)
quadraticYPredictionFourth <- beta0 + beta1*xVals + beta2*xVals^2
lines(xVals, quadraticYPredictionFourth, col = 'blue', lty=2)

```
```{r}
summary(quadraticRegressionFourth)
confint(quadraticRegressionFourth)
```
The R-squared value here is super small, which means that the quadratic model doesn't fit the data.


####In conclude, I think the best model is the multiple linear regression model.








