---
title: "class04_Lab Yixuan He"
output: html_document
---
## LAB for Class04

**The Goal**
Given a data set containing characteristics about the Boston housing market, explore methods for predicting housing prices.

You'll need to clone this repository using RStudio first. 
Practice writing and adding code that produces a simple linear regression, multiple linear regression, and polynomial regression.
Include a visual and quantitative method for assessing whether some models may better explain housing prices than other models.

We will discuss some models students have proposed and ways to communicate these models to a lay audience.

```{r}
library(readr)
library(dplyr)
library(ggplot2)

bostonHousing <- read.csv("data/boston_data.csv")
```

Simple Linear Regression
```{r}
slr_fit <- lm(MEDV~CRIM,data=bostonHousing)
summary(slr_fit)

plot(bostonHousing$CRIM,bostonHousing$MEDV,main="Boston Housing", 
     xlab="Crime Rate", ylab="Median Value")
abline(slr_fit)
plot(slr_fit)

sse_slr <- sum(resid(slr_fit)^2)
print(sse_slr)
```

Multiple Linear Regression
```{r}
mtp_fit1 <- lm(MEDV~CRIM+DIS+LSTAT,data=bostonHousing)
summary(mtp_fit1)

sse_mtp1 <- sum(resid(mtp_fit1)^2)
print(sse_mtp1)

plot(mtp_fit1)


mtp_fit2 <- lm(MEDV~CRIM+ZN+INDUS+CHAS+NOX+RM+AGE+DIS+RAD+TAX+B+LSTAT,data=bostonHousing)
summary(mtp_fit2)

sse_mtp2 <- sum(resid(mtp_fit2)^2)
print(sse_mtp2)

plot(mtp_fit2)
```

Polynomial Regression
```{r}
pln_fit1 <- lm(MEDV~poly(CRIM,2,raw=TRUE),data=bostonHousing)
summary(pln_fit1)

sse_pln1 <- sum(resid(pln_fit1)^2)
print(sse_pln1)

plot(pln_fit1)


pln_fit2 <- lm(MEDV~poly(CRIM,3,raw=TRUE),data=bostonHousing)
summary(pln_fit2)

sse_pln2 <- sum(resid(pln_fit2)^2)
print(sse_pln2)

plot(pln_fit2)
```


### Data
The dataset and the data dictionary are included in the folder called data.

### Questions

* How would you explain SSE?

SSE presents the difference between each data point and the mean value, which gives us an idea of how much variation there is in the data we choose. Simply looking at the value, we can see that the multiple linear regression with all variables gives us the least variation, indicating that it is the best model among the ones I have tried here.

* How would you explain your model?

I fitted five models in total here, one simple linear regression model examining the relationship between crime rate and housing value (slr_fit), one multiple linear regression model with variables that are most obviously negative correlated to housing value (mtp_fit1) and one with all independent variables (mtp_fit2), and two polynomial regression models using crime rate data to the degree of 2 (pln_fit1) and 3 (pln_fit2).

The slr_fit model does not work very well with high SSE, low R-squared value and not fitted line. The mtp_fit1 model is better with higher R_squared value and a more fitted and flatter residual plot. The mtp_fit2 model is the best among all, which makes intuitive sense since I included all independent variables. Both polynomial regression models have similar problems with the slr_fit model, but we can see that R-squared values are higher than the slr_fit model, which indicates they are a better prediction. If I could do this for all variables to higher degrees, I might be able to find a better model than the mtp_fit2 model.

### Resources
http://statweb.stanford.edu/~jtaylo/courses/stats203/R/multiple/multiple.R.html

